---
title: "Reference Interchange Format (RIF)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background

The Reference Interchange Format (RIF) is a simple text format for storing and exchanging growth references. Growth references typically consist of a conditioning variable, usually age or height, and a set of parameter estimates at each tabulated age or height. For example, the first four rows of the WHO Growth Standard for lengths of boys are

```{r setup}
library(centile)
ref <- load_reference("who_2006_hgt_male_")
head(ref, 4)
```

In this example, `x` stands for age in years and the `L`, `M` and `S` columns stand for the estimated parameters of the LMS model (Cole & Green 1992). The `study` attributes stores some additional information:

```{r attr}
head(attr(ref, "study"))
```

The RIF format is a simple way to store references as text files that can serve as input for different systems.

## General specification

Let us first take a look at an example. The following block contains the first 12 rows of the file `who_2006_bmi_female_.txt`.

```
name = who
year = 2006
yname = bmi
sex = female
distribution = LMS
citation = WHO 2006
publication = WHO Multicentre Growth Reference Study Group (2006). WHO Child Growth Standards: Length/height-for-age, weight-for-age, weight-for-length, weight-for-height and body mass index-for-age: Methods and development. Geneva: World Health Organization; pp 312. (web site: http://www.who.int/childgrowth/publications/en/)
[data]
x	L	M	S
0	-0.0631	13.3363	0.09272
0.0027	0.0362	13.3185	0.0936
0.0055	0.1355	13.3006	0.09448
...
```

A RIF formatted files should follow the following principles:

- A RIF file contains the parameters of one reference distribution only. Thus, separate distributions for males and females go into different files;
- The file starts with a header that describe the contents of the reference table. The section below provides the accepted keywords;
- The `=`  sign separates the keyword and its value;
- A special keyword `[DATA]` separates the header and data;
- The first line after `[DATA]` contain the variable names. These names adhere to certain naming conventions given below;
- There is one column named `x`, the conditioning variable;
- Data columns are separated by tabs.

## Keywords

Keyword       | Required    | Interpretation
------------- | ----------- | ---------------
name          | Yes         | two-letter country code, `who`, `eu27`, `fenton`, and so on
year          | Yes         | four number, may be left blank
yname         | Yes         | Outcome: `hgt`, `wgt`, `hdc`, `wfh`, `bmi`, `dsc`
sex           | No          | `male` or `female`, may be left blank
sub           | No          | Subgroup
distribution  | Yes         | `NO`, `LMS`, `BCCG`, `BCPE` or `BCT`
citation      | No          | Short citation phrase
publication   | No          | Reference to source documents
remark        | No          | Free text for additional remarks
tx            | No          | Transformation of `x` prior to Z-score calculation
xt            | No          | Back-transform to `x` after Z-score calculation
ty            | No          | Transformation of `y` prior to Z-score calculation
yt            | No          | Back-transform to `y` after Z-score calculation

Any other keywords, each on its own line, can be added and will be saved in `attr(ref, "study")`. The number of header rows is limited to 24.

## Column names

Apart from the column named `x` two or more column with parameters, depending on the specification of the distribution.

Distribution  | Required                    | Interpretation
------------- | --------------------------- | ---------------
NO            | `mean`, `sd`                | Mean, Standard deviation
LMS           | `M`, `S`, `L`               | Location, Scale, Skewness
BCCG          | `mu`, `sigma`, `nu`         | Location, Scale, Skewness
BCPE          | `mu`, `sigma`, `nu`, `tau`  | Location, Scale, Skewness, Kurtosis
BCT           | `mu`, `sigma`, `nu`, `tau`  | Location, Scale, Skewness, Kurtosis

Some other distributions can be specified, but no Z-scores will be calculated:

Distribution  | Required                    | Interpretation
------------- | --------------------------- | ---------------
MEA           | `mean`                      | Mean only
PCT           | `pxx`, `pyy`                | One or more percentiles (xx, yy numeric)
MP            | `mean`, `pxx`, `pyy`        | Mean + Percentiles (xx, yy numeric)
